Proposed Solution

### 2017-01-24 CAV17/secs/eval.tex
fxing learner

\begin{table}[h]
\centering
\caption{Results of \app}
\label{table:learning}
\setlength{\tabcolsep}{0.5em}
\begin{tabular}{|c|c|c|c|c|}
\hline
{\bf Class of Error } & {\bf \# Rules Learned} & {\bf \# Rules Detected} & {\bf Support} & {\bf Confidence}\\ 
\hline
\hline
Order        & 13  & 389  &  & \\ 
Missing      & 53  & 237  &  & \\ 
Type         & 92  & 1367 &  & \\ 
Fine-Grain   & 327 & 55   &  & \\ 
Coarse-Grain & 97  & 63   &  & \\ 
\hline 
\end{tabular}
\end{table}

@@

\begin{table}[h]
\centering
\caption{Effect of Post Analysis Rule Ordering for known true positives}
\label{table:order}
\setlength{\tabcolsep}{1em}
\begin{tabular}{|c|c|c|c|c|}
\hline
{\bf Class of Error} & {\bf Reference} & {\bf Rank w/o} & {\bf Rank w/} & {\bf Ideal Rank}\\
\hline
\hline
Order        & Ex.~\ref{ex:order}   & & \\ 
Missing      & Ex.~\ref{ex:miss}    & & \\ 
Type         & Ex.~\ref{ex:type}    & & \\ 
Fine-Grain   & Ex.~\ref{ex:fine}    & & \\ 
Coarse-Grain & Ex.~\ref{ex:coarse}  & & \\ 
\hline
\end{tabular}
\end{table}

The checker stage of \app is the only one facing an end user that wants to verify a new configuration file.
However, the user may also want to interface with the previous stages as well. 
We have only build rules for MySQL, a relatively well documented language, but any configuration language can be analyzed with \app given a training set.
In this case, the user will need to rebuild the rule set.
A user also may want to update the support of confidence threshold to report more or fewer error, as is needed for their application.
It is then important that the learning process is reasonably fast.
The times reported in Table~\ref{table:training} were run on a single thread of a Intel i5-3450 CPU @ 3.10GHz with 16GB of RAM.
We see the drastic difference in times here because ConfigC assumes a correct training set, and learns every relation derivable.
With \app, we instead only process rule that meet the required support and confidence, reducing the cost of resolve to a consistent set of rules.

\begin{table}[h]
\centering
\caption{Training time for different datasets}
\label{table:training}
\setlength{\tabcolsep}{1em}
\begin{tabular}{|c|c|c|}
\hline
{\bf The \# of Files for Training} & {\bf ConfigC (s)} & {\bf \app (s)}\\ 
\hline
\hline
10   & 2.997     &   \\ \hline
50   & 46.4208   &   \\ \hline
100  & 206.7798  &   \\ \hline
150  & 431.7772  &   \\ \hline
200  & 856.6383  &   \\ 
\hline
\end{tabular}
\end{table}

>>>>

\begin{table}[h]
\centering
\caption{Results of \app}
\label{table:learning}
\setlength{\tabcolsep}{0.5em}
\begin{tabular}{|c|c|c|c|c|}
\hline
{\bf Class of Error } & {\bf \# Rules Learned} & {\bf \# Errors Detected} & {\bf Support} & {\bf Confidence}\\ 
\hline
\hline
Order        & 13  & 389  &  & \\ 
Missing      & 53  & 237  &  & \\ 
Type         & 92  & 1367 &  & \\ 
Fine-Grain   & 327 & 55   &  & \\ 
Coarse-Grain & 97  & 63   &  & \\ 
\hline 
\end{tabular}
\end{table}

@@

\begin{table}[h]
\centering
\caption{Effect of Post Analysis Rule Ordering for known true positives}
\label{table:order}
\setlength{\tabcolsep}{1em}
\begin{tabular}{|c|c|c|c|}
\hline
{\bf Class of Error} & {\bf Reference} & {\bf Rank w/o} & {\bf Rank w/}\\
\hline
\hline
Order        & Ex.~\ref{ex:order}   &  & \\ 
Missing      & Ex.~\ref{ex:miss}    &  & \\ 
Type         & Ex.~\ref{ex:type}    &  & \\ 
Fine-Grain   & Ex.~\ref{ex:fine}    &  & \\ 
Coarse-Grain & Ex.~\ref{ex:coarse}  &  & \\ 
\hline
\end{tabular}
\end{table}

The checker stage of \app is the only one facing an end user that wants to verify a new configuration file.
Once a set of rules has been learned, it is not necessary to rerun the learner.
However, we have only used \app to build rules for MySQL, but any configuration language can be analyzed with \app given a training set.
In an industrial setting, the available training set may be much larger than ours as well, so is important that the learning process scales.
We see in Table~\ref{table:training} that \app scales roughly linearly.

\markk{TODO rewrite}
We also compare \app to prior work in configuration verification, ConfigC~\cite{santolucitoCAV}.
ConfigC scales exponentially because the learning algorithm assumes a completely correct training set, and learns every derivable relation.
With \app, we instead only process rules that meet the required support and confidence, reducing the cost of resolving to a consistent set of rules. 
The times reported in Table~\ref{table:training} were run on four cores of a Kaby Lake Intel Core i7-7500U CPU @ 2.70GHz on 16GB RAM and Fedora 25.

\begin{table}[h]
\centering
\caption{Time for training over various training set sizes}
\label{table:training}
\setlength{\tabcolsep}{1em}
\begin{tabular}{|c|c|c|}
\hline
{\bf \# of Files for Training} & {\bf ConfigC (sec)} & {\bf \app (sec)}\\ 
\hline
\hline
0    & 0.051    & 0.51  \\ \hline
50   & 1.815    & 1.638  \\ \hline
100  & 13.331   & 4.119  \\ \hline
150  & 95.547   & 4.231  \\ \hline
200  & 192.882  & 12.271  \\ \hline
256  & 766.904  & 15.627  \\ 
\hline
\end{tabular}
\end{table}

### 2017-01-24 CAV17/secs/intro.tex
fxing learner

>>>>

\markk{maybe we can use these lines in the intro somewhere.}
These errors can cause total system failures, but can also be more insidious, for example slowing down the system only when the server load increases beyond a certain threshold.
Since these runtime errors may only be triggered after some time in a deployment environment, the standard delta-debugging technique~\cite{x} of starting a system multiple times with different configuration settings will not help detect these misconfigurations.

### 2017-01-24 CAV17/secs/learner.tex
fix fine grain
fxing learner

The goal of the learner is to derive rules from the intermediate representation of the training set generated by the translator.
We describe an interface to define the different classes of rules that should be learned.
Each instance of the interface corresponds to a different class of of configuration errors, such as missing entry errors, ordering errors, and integer correlation errors. 
These errors can cause total system failures, but can also be more insidious, for example slowing down the system only when the server load increases beyond a certain threshold.

\rahul{and we may even be able to say that these insidiuous issues are much
harder to pin down by the standard delta-debugging technique of starting a
system multiple times with different configuration settings}

A rule is an implication relationship, $X \implies p(X,Y)$, between two possibly empty sets of keywords $X,Y \subset \keys$, where $\keys$ is the set of unique keys from the training set and the predicate $p$ is the one of the classes of configuration errors (order, missing, etc).
Implicitly we interpret this to means that if the keywords $X\cup Y$ appear in a configuration file, we expect the predicate $p$ to hold.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
These two key metrics are taken from \textit{association rule learning}~\cite{agrawal1993mining}, a technique that can be summarized as inductive machine learning.
In the configuration verification domain, standard association rule learning is best suited to learn integer correlation and missing keyword rules.
In fact, a more specialized technique for learning ordering rules would be sequence mining~\cite{}, and again another approach may be a better fit for rules over a single keyword, as is needed for probabilistic types.
Since the full algorithm details are out-of-scope of this paper, here we only describe in detail the metrics which are most relevant to verification.

%TODO math def of support and confidence
Each power set of keywords, $\{X,Y\}$, is assigned a support and confidence measure during the learning process.
Support is the number of times the set of keywords in the proposed rule have been seen the in the training set.
Confidence is the number of times the rule predicate has held true over the given keywords.
In the learning process, these are given threshold, below which a rule will be reject for lack of evidence.

\para{Order}
Ordering errors take the form $X \implies order(X,Y)$, where $|X|,|Y|=1$ and the keyword $X$ must come before the keyword $Y$ in any configuration file.

\para{Missing}
$X \implies missing(X,Y)$, where $|X|,|Y|=1$ and the keyword $X$ must appear in the same file as the keyword $Y$ in any configuration file.

@@

We have developed \app to be easily extendable and customizable. Each of the rules are implemented as an instance of the \textit{Learnable} typeclass:

@@

With the learned rules generates by the learner module,
  \app checks whether any entry in a target configuration file violates the learned rules and constraints.
\app parses a verification target configuration file the same way employed in the translator for learning,
  obtaining a structured and typed representation.
Then, the checker applies the learner to build the set of relations observed in the file.
For any relation that violates a known rule, the checker will output the predicate and keyword sets associated with that rule, as well as the learned support and confidence values.
Since the tool is probabilistic, we provide the user with these values to determine if they rule must be satisfied in their application on a case-by-case basis.
For instance, the \texttt{key\_buffer} misconfiguration from Sec. \ref{ex:fine} will only be noticeable if the application experiences a heavy traffic load, so the user may choose to ignore this error if they are confident this will not be an issue.

>>>>

The goal of the learner is to derive rules from the intermediate representation of the training set generated by the translator.
We describe an interface to define the different classes of rules that should be learned.
Each instance of the interface corresponds to a different class of configuration errors, as described in Sec.~\ref{sec:motiv}.

To learn rules over sets of configuration files, we use a generalization of \textit{association rule learning}~\cite{agrawal1993mining}, a technique that can be summarized as inductive machine learning.
Association rule learning is a technique to learn how frequently items of a set appear together.
For example, by examining a list of food store receipts, we would learn that when a customer buys bread and peanut butter, the set of purchased items is also likely to include jelly.
Since configuration files have complex relations, we extend these association relationships to generalized predicates.

A \textit{rule}, $r$, is then an implication relationship, $r = X \implies p(X,Y)$, between two possibly empty sets of keywords $X,Y \subset \keys$, where $\keys$ is the set of unique keys from the training set (denoted \trainingSet) and the predicate $p$ is one of the classes of configuration errors.
Implicitly we interpret a rule to mean that if the keywords $X\cup Y$ appear in a configuration file, the predicate $p$ should hold.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
The set of rules learned from training set \trainingSet constitutes a specification, $Spec(\trainingSet)$, for a configuration file to be considered correct.

%TODO math def of support and confidence
The two metrics support and confidence are used in association rule learning, as well as other rule based machine learning techniques~\cite{han2007frequent,langley1995applications}.
We use slightly modified definitions of support and confidence to handle arbitrary predicates as rules.
Each rule is assigned a support and confidence to measure the truth of the rule during the learning process.
%
\begin{equation}
 support(p,X,Y) = \frac{|\{C \in \trainingSet \mid X \cup Y \subseteq C\}|} {|\trainingSet|}
\end{equation}
\begin{equation}
 confidence(p,X,Y) = \frac{|\{C \in \trainingSet \mid p(X,Y) \subseteq C\}| } {support(X,Y)*|\trainingSet|}
\end{equation}

Support is the frequency that the that set of keywords in the proposed rule, $X \cup Y$, have been seen in the configuration files $C$ in the training set \trainingSet.
Notice support does not consider the predicate itself, only how many times the predicate had the chance to be evaluated.
Confidence is the number of times the rule predicate has held true over the given keywords.
In the learning process, each class of rule is manually assigned a support and confidence threshold, below which a rule will be rejected for lack of evidence.
The set of learned rules is then:

\begin{equation}
  Spec(\trainingSet) = \forall X,Y \subset \keys. {r | support(p,X,Y) > t_s \land confidence(p,X,Y) > t_c}
\end{equation}

\subsection{Error Classes}
Ordering errors take the form $X \implies order(X,Y)$, where $|X|,|Y|=1$ and the keyword $X$ must come before the keyword $Y$ in any configuration file.
Missing keyword entries take the form $X \implies missing(X,Y)$, where $|X|,|Y|=1$ and the keyword $X$ must appear in the same file as the keyword $Y$ in any configuration file.

@@

We have developed \app to be easily extendable and customizable. Each of the rules are implemented as an instance of the \textit{Learnable} interface (a typeclass in Haskell):

@@

With the rules generated by the learner module, \app checks whether any entry in a target configuration file violates the learned rules and constraints.
\app parses a verification target configuration file the same way employed in the translator for learning to obtain a set of key-value pairs $C$.
Then, the checker applies the learner to build the set of relations observed in the file.
For any relation that violates a known rule, the checker will output the predicate and keyword sets associated with that rule, as well as the learned support and confidence values.
Since the tool is probabilistic, we provide the user with these values to determine if they rule must be satisfied in their application on a case-by-case basis.
For instance, the \texttt{key\_buffer} misconfiguration from Sec. \ref{ex:fine} will only be noticeable if the application experiences a heavy traffic load, so the user may choose to ignore this error if they are confident this will not be an issue.
